Wireless Automobile Detection, License Plate Processing, and Data Availability Network Proposal

### 2014-03-12 proposal/proposal.tex
Progress

\subsection{Related Work}
The ACMx group at CSM has designed a system called SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are Arduino Fio microcontorllers equipped with tri-axial magnetometors and XBEE Pro radios. The base station is one small, commercially available Raspberry Pi Linux computer. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Fio and the Raspberry Pi. This system uses the ubiquitous IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing this proposal, the group has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information about project progress. The project described in this proposal will augment and extend the ACMx project while collaborating closely with Stillwell (ACMx president and author of \cite{stillwell2013}), with the goal of a real world deployment by May 2014.

\section{Proposed Work}
\subsection{Overview}
The project we are proposing will be an extension of the ACMx SmartLots project. We will collaborate closely with the ACMx group to create and deploy the first working parking lot monitoring system on the Mines campus. Our system will consist of several subsystems, each of which will be either an extension of a current ACMx subsystem or a new subsystem that will interface with the ACMx architecture. Figure \ref{fig:system} shows the overall architecture of our system, overlaid upon a non-scaled map of the CTLM parking lots.

@@

The system will rely on three sensor nodes: one will be placed at the entrance to the upper lot, one at the entrance to the lower lot, and one at the junction between the two lots. Each sensor node will contain an Arduino Fio equipped with a tri-axial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by ACMx. Our contribution will be to also equip each node with a Raspberry Pi mini-computer and camera, which will take pictures of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send an interrupt to the Raspberry Pi via the Xbee radio, waking it up so it can take a picture of the license plate. There will be a central base station in an upper window of the CTLM building, which will receive all the data from the sensor nodes via transmission from the Xbee Pro radios attached to the sensor nodes. This base station Raspberry Pi will also have a camera attached that will take periodic images of the lots from above. These images will be used to reconcile the sensed data with the actual state of the lots. The base station will forward the data to the ACMx server, where it will be processed and displayed in the web application. The following sections describe the architecture and function of each of the aforementioned subsystems in detail.

@@

The use of a magnetometer ensures that solely automobile and other large vehicles such as motorcycles are detected while pedestrians and particulate deposits, such as snow, are ignored.

\subsubsection{Tests and Analyses}
Work on the automobile detection subsystem will involve a variety of tests and analyses to ensure system and data integrity.

>>>>

\subsection{Related Work}
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and Xbee Pro radios. The base station is one small, commercially available Raspberry Pi Linux computer. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Fio and the Raspberry Pi. This system uses the ubiquitous IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing this proposal, the group has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information about project progress. The project described in this proposal will augment and extend the ACMx project while collaborating closely with Stillwell (ACMx president and author of \cite{stillwell2013}), with the goal of a real world deployment by May 2014.

\section{Proposed Work}
\subsection{Overview}
The project we are proposing will be an extension of the ACMx SmartLots project. We will collaborate closely with Stillwell (author of \cite{stillwell2013}) to create and deploy the first working parking lot monitoring system on the Colorado School of Mines campus. Our system will consist of several components, each of which will be either an extension of a current SmartLots system or an entirely new subsystem that will be cohesively interfaced. Figure \ref{fig:system} shows the overall architecture of our system.

@@

The system will rely on three sensor nodes: one will be placed at the entrance to the upper lot, one at the entrance to the lower lot, and one at the junction between the two lots. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by ACMx. Our contribution will be to also equip each node with a Raspberry Pi mini-computer and camera, which will take pictures of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send an interrupt to the Raspberry Pi via the Xbee radio, waking it up so it can take a picture of the license plate. There will be a central base station in an upper window of the CTLM building, which will receive all the data from the sensor nodes via transmission from the Xbee Pro radios attached to the sensor nodes. This base station Raspberry Pi will also have a camera attached that will take periodic images of the lots from above. These images will be used to reconcile the sensed data with the actual state of the lots. The base station will forward the data to the ACMx server, where it will be processed and displayed in the web application. The following sections describe the architecture and function of each of the aforementioned subsystems in detail.

@@

The use of a magnetometer ensures that solely automobiles and other roadway vehicles which such as motorcycles are detected while pedestrians and particulate deposits, such as snow, are ignored. As in \cite{stillwell2013}, we shall use the commercially available HMC5883 triaxial magnetometer as it provides an adequate balance of sensitivity and low power consumption.

\subsubsection{Tests and Analyses}
Work on the automobile detection subsystem will involve a variety of tests and analyses to ensure system and data integrity. Power consumption analyses will be conducted on the automobile detection and license plate image acquisition subsystem to identify power usage per main component (Figure~\ref{fig:autodetect}) while the mote is in different states, such as sleeping and detecting vehicles. Such an analysis will allow us to allocate how much time the mote is allowed to spend in each state and determine the required lithium polymer battery capacity.

Analyses will also be performed on the timing of different actions within the network, including:
\begin{itemize}
\item Automobile detection lag
\item Imaging of vehicle
\item License plate recognition and extraction
\item Packet transmission
\end{itemize}
This will help determine wether certain portions of the image data processing may need to be performed on a server and wether it would be feasible to funnel the image data through the Xbee instead of a dedicated directional IEEE 802.11 system.

The efficacy of the automobile detection subsystem will also be evaluated through several tests. A variety of different sized vehicles will be passed through the system at different velocities to ensure high detection accuracy. We will also analyze the end results of deployment after a certain period of time, to determine how often the automobile count needs to be reconciled.
